CrypTool Script \\ Mathematics and Cryptography

### 2003-03-31 script/de/chapters/primes.tex
be update fixes

\\
Auch wenn sich Fermat bei seiner Vermutung irrte, so stammt in diesem
Zusammenhang von ihm doch ein sehr wichtiger Satz: Der (kleine)
Fermatsche Satz, den Fermat im Jahr 1640 aufstellte, ist der 
Ausgangspunkt vieler schneller Primzahltests 
(\hyperlink{KleinerSatzFermat-chap3}{siehe Kap. 3.8.3}).

@@

    Innerhalb des Projektes ''Distributed Search for Fermat Number Dividers'',
    das von Leonid Durman angeboten wird, gibt es ebenfalls Fortschritte 
    beim Finden von neuen Primzahl-Riesen
    ({\href{http://www.fermatsearch.org/}
       {\tt http://www.fermatsearch.org/}}  --  diese Webseite hat 
    Verkn"upfungen zu Seiten in russisch, italienisch und deutsch).

    Die entdeckten Faktoren k"onnen sowohl zusammengesetzte nat"urliche als auch 
    prime nat"urliche Zahlen sein.

@@

    Mit diesem Programm fand Michael Angel am 16. Februar 2003 eine
    prime verallgemeinerte Fermatzahl mit 628.808 Dezimalstellen, 
    die zum damaligen Zeitpunkt zur 5.-gr"o"sten bis dahin bekannten
    Primzahl wurde:
    $$ b^{2^{17}} + 1  =  62.722^{131.072} + 1. $$ 
    Dadurch wurde die erste im GIMPS-Projekt\index{GIMPS} gefundene
    Mersenne-Primzahl M-35 = M1398269 auf den 6. Platz verwiesen.

@@

\item \index{Euklidzahlen} {\em Euklidzahlen} $e_n = e_0 \cdot e_1 \cdots 
    e_{n-1} + 1$  mit $n$ gr"o"ser oder gleich $1$ und $e_0 := 1$. \\
    $e_{n-1}$ ist nicht die $(n-1)$-te Primzahl, sondern die zuvor hier 
    gefundene Zahl.
    Diese Formel ist leider nicht offen, sondern rekursiv.
    Die Folge startet mit
$$
\small\begin{array}{lll}
e_1 = 1 + 1 &   = 2 &   \mapsto {\rm ~prim} \\
e_2 = e_1 + 1   &   = 3 &   \mapsto {\rm ~prim} \\
e_3 = e_1 \cdot e_2 + 1 &   = 7 &   \mapsto {\rm ~prim} \\
e_4 = e_1 \cdot e_2 \cdot e_3 + 1 & = 43 &  \mapsto {\rm ~prim} \\
e_5 = e_1 \cdot e_2 \cdots e_4 + 1 &    = 13 \cdot 139 &    \mapsto {\rm ~NICHT~prim} ! \\
e_6 = e_1 \cdot e_2 \cdots e_5 + 1 &    = 3.263.443 &   \mapsto {\rm ~prim} \\
e_7 = e_1 \cdot e_2 \cdots e_6 + 1 &    = 547 \cdot 607 \cdot 1.033 \cdot 31.051 & \mapsto {\rm ~NICHT~prim} ! \\
e_8 = e_1 \cdot e_2 \cdots e_7 + 1 &    = 29.881\cdot 67.003 \cdot 9.119.521 \cdot 6.212.157.481 & \mapsto {\rm ~NICHT~prim} !
\end{array}
$$
Auch $e_9, \cdots, e_{17}$ sind zusammengesetzt, so dass dies auch
keine brauchbare Formel ist. Bemerkung: Das Besondere an 
diesen Zahlen ist, dass sie jeweils paarweise keinen
gemeinsamen Teiler au"ser $1$ haben, sie sind also%
\index{Primzahlen!relative}\index{relativ prim} {\em relativ zueinander prim}.

@@

% --------------------------------------------------------------------------
\newpage
\subsubsection{Gr"o"senordnungen / Dimensionen in der Realit"at}
\label{s:grosord}
Bei der Beschreibung kryptographischer Protokolle und Algorithmen
treten Zahlen auf, die so gro"s bzw. so klein sind, dass sie einem
intuitiven Verst"andnis nicht zug"anglich sind. Es kann daher
n"utzlich sein, Vergleichszahlen aus der uns umgebenden realen
Welt bereitzustellen, so dass man ein Gef"uhl f"ur die Sicherheit
kryptographischer Algorithmen entwickeln kann. Die angegebenen
Werte stammen teilweise aus \cite{Schwenk1996} und
\cite[S.18]{Schneier1996}.  \hypertarget{grosord}{}
\begin{tabbing}
Wahrscheinlichkeit, dass Sie  auf ihrem n"achsten Flug entf"uhrt
werden:~~ \= abcdefhijk \= \kill
Wahrscheinlichkeit, dass Sie  auf ihrem n"achsten Flug entf"uhrt werden \> $ 5,5 \cdot 10^{-6} $\> \\
Wahrscheinlichkeit f"ur 6 Richtige im Lotto \> $ 7,1 \cdot 10^{-8} $\> \\
J"ahrliche Wahrscheinlichkeit, von einem Blitz getroffen zu werden \> $ 10^{-7} $\> \\
Risiko, von einem Meteoriten erschlagen zu werden \> $ 1,6 \cdot 10^{-12} $ \> \\
------------------------------------------------------------------------------------------------------------------\\
Zeit bis zur n"achsten Eiszeit (in Jahren) \> $14000 $ \> $ (2^{14})$ \\
Zeit bis die Sonne vergl"uht (in Jahren)   \> $10^{9} $ \> $(2^{30})$ \\
Alter der Erde (in Jahren)\> $ 10^9 $ \> $(2^{30}) $  \\
Alter des Universums (in Jahren) \> $ 10^{10} $ \> $(2^{34}) $ \\
Anzahl der Atome der Erde              \> $10^{51} $ \> $ (2^{170}) $ \\
Anzahl der Atome der Sonne             \> $10^{57}$ \> $ (2^{190})$ \\
Anzahl der Atome im Universum (ohne dunkle Materie)            \> $10^{77}$  \> $ (2^{265})$ \\
Volumen des Universums (in $cm^3$)     \> $10^{84}$ \> $(2^{280})$
\end{tabbing}

@@

%already defined in elementaryNumberTheory.inc -> 2 davor
\bibitem[Schneier1996]{2Schneier1996} \index{Schneier 1996}
    Bruce Schneier, \\
    {\em Applied Cryptography, Protocols, Algorithms, and Source Code in C},\\
    Wiley and Sons, 2nd edition 1996.

>>>>

Auch wenn sich Fermat bei seiner Vermutung irrte, so stammt in diesem
Zusammenhang von ihm doch ein sehr wichtiger Satz: Der (kleine)
Fermatsche Satz, den Fermat im Jahr 1640 aufstellte, ist der 
Ausgangspunkt vieler schneller Primzahltests 
(\hyperlink{KleinerSatzFermat-chap3}{siehe Kap. 3.8.3}).

@@

    Innerhalb des Projektes ''Distributed Search for Fermat Number Dividers'',
    das von Leonid Durman angeboten wird, gibt es ebenfalls Fortschritte 
    beim Finden von neuen Primzahl-Riesen
    ({\href{http://www.fermatsearch.org/}
       {\tt http://www.fermatsearch.org/}}  --  diese Webseite hat 
    Verkn"upfungen zu Seiten in russisch, italienisch und deutsch).
    \begin{sloppypar}
    Die entdeckten Faktoren k"onnen sowohl zusammengesetzte nat"ur"-liche als auch 
    prime nat"urliche Zahlen sein.
    \end{sloppypar}

@@

    Mit diesem Programm fand Michael Angel am 16. Februar 2003 eine
    prime verallgemeinerte Fermatzahl mit 628.808 Dezimalstellen, 
    die zum damaligen Zeitpunkt zur 5.-gr"o"sten bis dahin bekannten
    Primzahl wurde:
    $$ b^{2^{17}} + 1  =  62.722^{131.072} + 1. $$ 
    Dadurch wurde die erste im GIMPS-Projekt\index{GIMPS} gefundene
    Mersenne-Primzahl $$\mbox{M-35 = M1398269}$$ auf den 6. Platz verwiesen.

@@

\item \index{Euklidzahlen} {\em Euklidzahlen} $e_n = e_0 \cdot e_1 \cdots 
    e_{n-1} + 1$  mit $n$ gr"o"ser oder gleich $1$ und $e_0 := 1$. \\
    $e_{n-1}$ ist nicht die $(n-1)$-te Primzahl, sondern die zuvor hier 
    gefundene Zahl.
    Diese Formel ist leider nicht offen, sondern rekursiv.
    Die Folge startet mit
$$
\begin{array}{lll}
e_1 = 1 + 1 &   = 2 &   \mapsto {\rm ~prim} \\
e_2 = e_1 + 1   &   = 3 &   \mapsto {\rm ~prim} \\
e_3 = e_1 \cdot e_2 + 1 &   = 7 &   \mapsto {\rm ~prim} \\
e_4 = e_1 \cdot e_2 \cdot e_3 + 1 & = 43 &  \mapsto {\rm ~prim} \\
e_5 = e_1 \cdot e_2 \cdots e_4 + 1 &    = 13 \cdot 139 &    \mapsto {\rm ~NICHT~prim} ! \\
e_6 = e_1 \cdot e_2 \cdots e_5 + 1 &    = 3.263.443 &   \mapsto {\rm ~prim} \\
e_7 = e_1 \cdot e_2 \cdots e_6 + 1 &    = 547 \cdot 607 \cdot 1.033 \cdot 31.051 & \mapsto {\rm ~NICHT~prim} ! \\
e_8 = e_1 \cdot e_2 \cdots e_7 + 1 &    = 29.881\cdot 67.003 \cdot 9.119.521 \cdot 6.212.157.481 & \mapsto {\rm ~NICHT~prim} !
\end{array}
$$
Auch $e_9, \cdots, e_{17}$ sind zusammengesetzt, so dass dies auch
keine brauchbare Formel ist. Bemerkung: Das Besondere an 
diesen Zahlen ist, dass sie jeweils paarweise keinen
gemeinsamen Teiler au"ser $1$ haben, sie sind also%
\index{Primzahlen!relative}\index{relativ prim} {\em relativ zueinander prim}.

@@

% --------------------------------------------------------------------------
\newpage
\subsubsection{Gr"o"senordnungen / Dimensionen in der Realit"at}
\label{s:grosord}
Bei der Beschreibung kryptographischer Protokolle und Algorithmen
treten Zahlen auf, die so gro"s bzw. so klein sind, dass sie einem
intuitiven Verst"andnis nicht zug"anglich sind. Es kann daher
n"utzlich sein, Vergleichszahlen aus der uns umgebenden realen
Welt bereitzustellen, so dass man ein Gef"uhl f"ur die Sicherheit
kryptographischer Algorithmen entwickeln kann. Die angegebenen
Werte stammen teilweise aus \cite{Schwenk1996} und
\cite[S.18]{Schneier1996p}.  \hypertarget{grosord}{}
\begin{tabbing}
Wahrscheinlichkeit, dass Sie  auf ihrem n"achsten Flug entf"uhrt
werden:~~ \= abcdefhijk \= \kill
Wahrscheinlichkeit, dass Sie  auf ihrem n"achsten Flug entf"uhrt werden \> $ 5,5 \cdot 10^{-6} $\> \\
Wahrscheinlichkeit f"ur 6 Richtige im Lotto \> $ 7,1 \cdot 10^{-8} $\> \\
J"ahrliche Wahrscheinlichkeit, von einem Blitz getroffen zu werden \> $ 10^{-7} $\> \\
Risiko, von einem Meteoriten erschlagen zu werden \> $ 1,6 \cdot 10^{-12} $ \> \\
------------------------------------------------------------------------------------------------------------------\\
Zeit bis zur n"achsten Eiszeit (in Jahren) \> $14000 $ \> $ (2^{14})$ \\
Zeit bis die Sonne vergl"uht (in Jahren)   \> $10^{9} $ \> $(2^{30})$ \\
Alter der Erde (in Jahren)\> $ 10^9 $ \> $(2^{30}) $  \\
Alter des Universums (in Jahren) \> $ 10^{10} $ \> $(2^{34}) $ \\
Anzahl der Atome der Erde              \> $10^{51} $ \> $ (2^{170}) $ \\
Anzahl der Atome der Sonne             \> $10^{57}$ \> $ (2^{190})$ \\
Anzahl der Atome im Universum (ohne dunkle Materie)            \> $10^{77}$  \> $ (2^{265})$ \\
Volumen des Universums (in $cm^3$)     \> $10^{84}$ \> $(2^{280})$
\end{tabbing}

@@

%already defined in elementaryNumberTheory.inc -> 2 davor
\bibitem[Schneier1996]{Schneier1996p} \index{Schneier 1996}
    Bruce Schneier, \\
    {\em Applied Cryptography, Protocols, Algorithms, and Source Code in C},\\
    Wiley and Sons, 2nd edition 1996.

### 2003-03-31 script/en/chapters/cryptomethods.tex
be update fixes

Brute-force (exhaustive search, trial-and-error) means to completely examine all keys of the key space: so no special analysis methods have to be used. Instead, the cipher text is decrypted with all possible keys and for each resulting text it is checked, whether this is a meaningful clear text. A key length of 64 bit means at most 2^64 = 18,446,744,073,709,551,616 or about 18 trillion (GB) / 18 quintillion (US)  keys to check\index{CrypTool}%
\footnote{%
    With CrypTool\index{CrypTool} v1.3 you also can try brute-force attacks
    of modern symmetric algorithms 
    (using the menu {\bf Analysis \textbackslash{} Ciphertext-Only}): 
    to achieve a result in a appropriate time with a single PC you should 
    mark not more than 20 bit of the key as unknown.
}.

Companies like RSA Security provide so-called cipher challenges in order to quantify the security offered by well-known symmetric ciphers as DES, Triple-DES or RC5 (\href{http://www.rsasecurity.com/rsalabs/challenges/secretkey/index.html}{\tt http://www.rsasecurity.com/rsalabs/challenges/secretkey/index.html}). They offer prizes for those who manage to decipher cipher texts, encrypted with different algorithms and different key lengths and unveil the symmetric key (under controlled conditions). So theoretical estimates can be confirmed.

It is well-known, that the "old" standard algorithm DES with a fixed key length of 56 bit is no more secure: this was demonstrated already in January 1999 by the Electronic Frontier Foundation (EFF). With their specialized computer Deep Crack they cracked a DES encrypted message within less than a day (\href{http://www.rsasecurity.com/rsalabs/challenges/des3/index.html}{\tt http://www.rsasecurity.com/rsalabs/challenges/des3/index.html}).

@@

The RC5-64 challenge has been solved by the distributed.net team after 5 years (\href{http://distributed.net/pressroom/news-20020926.html}{\tt http://distributed.net/pressroom/news-20020926.html}).  At all, 331,252 individuals co-operated over the internet to find the key. More than 15 trillion (GB) / 15 quintillion (US)  keys were checked, until they found the right key.

@@

If Alice\index{Alice}%
\footnote{%
      In order to describe cryptographic protocols the participants
      are often named Alice, Bob, \cdots (see \cite[S. 23]{Schneier1996}). 
      Alice and Bob perform all 
      2-person-protocols. Alice will initiate all protocols and 
      Bob answers. The attackers are named Eve (eavesdropper) and
      Mallory (malicious active attacker).
  } wants to communicate with Bob, then she finds Bob's public key 
in the directory and uses it to encrypt her message to him. She then sends
this cipher text to Bob, who is then able to decrypt it again using his 
secret key. As only Bob knows his secret key, only he can decrypt 
messages addressed to him.
Even Alice who sends the message cannot restore plaintext from the (encrypted)
message she has sent. Of course, you must first ensure that the public key
cannot be used to derive the private key.\par \vskip + 3pt

@@

  \item distributed.net:~\glqq RC5-64 has been solved'' \\
        \href{http://distributed.net/pressroom/news-20020926.html}
         {\tt http://distributed.net/pressroom/news-20020926.html}

  \item RSA: ~\glqq The RSA Secret Key Challenge'' \\
       \href{http://www.rsasecurity.com/rsalabs/challenges/secretkey/index.html}
        {\tt http://www.rsasecurity.com/rsalabs/challenges/secretkey/index.html}

  \item RSA: ~\glqq DES Challenge'' \\
        \href{http://www.rsasecurity.com/rsalabs/challenges/des3/index.html}
         {\tt http://www.rsasecurity.com/rsalabs/challenges/des3/index.html}

>>>>

Brute-force (exhaustive search, trial-and-error) means to completely examine all keys of the key space: so no special analysis methods have to be used. Instead, the cipher text is decrypted with all possible keys and for each resulting text it is checked, whether this is a meaningful clear text. A key length of 64 bit means at most $2^{64}$ = 18,446,744,073,709,551,616 or about 18 trillion (GB) / 18 quintillion (US)  keys to check\index{CrypTool}%
\footnote{%
    With CrypTool\index{CrypTool} v1.3 you also can try brute-force attacks
    of modern symmetric algorithms 
    (using the menu {\bf Analysis \textbackslash{} Ciphertext-Only}): 
    to achieve a result in a appropriate time with a single PC you should 
    mark not more than 20 bit of the key as unknown.
}.

Companies like RSA Security provide so-called cipher challenges in order to quantify the security offered by well-known symmetric ciphers as DES, Triple-DES or RC5\footnote{\href{http://www.rsasecurity.com/rsalabs/challenges/secretkey/index.html}{\tt http://www.rsasecurity.com/rsalabs/challenges/secretkey/index.html}}. They offer prizes for those who manage to decipher cipher texts, encrypted with different algorithms and different key lengths and unveil the symmetric key (under controlled conditions). So theoretical estimates can be confirmed.

It is well-known, that the "old" standard algorithm DES with a fixed key length of 56 bit is no more secure: this was demonstrated already in January 1999 by the Electronic Frontier Foundation (EFF). With their specialized computer Deep Crack they cracked a DES encrypted message within less than a day\footnote{\href{http://www.rsasecurity.com/rsalabs/challenges/des3/index.html}{\tt http://www.rsasecurity.com/rsalabs/challenges/des3/index.html}}.

@@

The RC5-64 challenge has been solved by the distributed.net team after 5 years\footnote{\href{http://distributed.net/pressroom/news-20020926.html}{\tt http://distributed.net/pressroom/news-20020926.html}}.  In total 331,252 individuals co-operated over the internet to find the key. More than 15 trillion (GB) / 15 quintillion (US)  keys were checked, until they found the right key.

@@

If Alice\index{Alice}%
\footnote{%
      In order to describe cryptographic protocols the participants
      are often named Alice, Bob, \dots (see \cite[S. 23]{Schneier1996}). 
      Alice and Bob perform all 
      2-person-protocols. Alice will initiate all protocols and 
      Bob answers. The attackers are named Eve (eavesdropper) and
      Mallory (malicious active attacker).
  } wants to communicate with Bob, then she finds Bob's public key 
in the directory and uses it to encrypt her message to him. She then sends
this cipher text to Bob, who is then able to decrypt it again using his 
secret key. As only Bob knows his secret key, only he can decrypt 
messages addressed to him.
Even Alice who sends the message cannot restore plaintext from the (encrypted)
message she has sent. Of course, you must first ensure that the public key
cannot be used to derive the private key.\par \vskip + 3pt

@@

  \item distributed.net: ``RC5-64 has been solved'' \\
        \href{http://distributed.net/pressroom/news-20020926.html}
         {\tt http://distributed.net/pressroom/news-20020926.html}

  \item RSA: ``The RSA Secret Key Challenge'' \\
       \href{http://www.rsasecurity.com/rsalabs/challenges/secretkey/index.html}
        {\tt http://www.rsasecurity.com/rsalabs/challenges/secretkey/index.html}

  \item RSA: ``DES Challenge'' \\
        \href{http://www.rsasecurity.com/rsalabs/challenges/des3/index.html}
         {\tt http://www.rsasecurity.com/rsalabs/challenges/des3/index.html}

### 2003-03-31 script/en/chapters/numbertheory.tex
be update fixes

% --------------------------------------------------------------------------
\subsubsection{Patterns and Structures}
\hypertarget{kap_3_5_2}
In general mathematicians investigate ``Structures''\index{Structure}. They ask e.g. at $ a * x \equiv b $ mod $m$,
which values $x$ can take for given values of $a, ~b, ~m.$

@@

% --------------------------------------------------------------------------
\subsubsection{Patterns and Structures}
\hypertarget{Kap_3_8}{}

@@

This record, established on January 18, 2002 by researchers of the German
university of Bonn (see: \\
\href{http://www.ercim.org/publication/Ercim\_News/enw49/franke.html, 2002-01}
{\texttt{http://www.ercim.org/publication/Ercim\_News/enw49/franke.html, 2002-01}}), got much less attention within the press than the
solution of the RSA challenge, established on August 22, 1999, when
researchers from the Netherlands factorized the 155-digit number into its
both 78-digit primes (see Chapter \ref{chptSecurityParam}). 
One reason is that 512 bit RSA-155 was a kind of {\em magic} border.

The task of the researchers from Bonn was not initiated by a challenge, but
they wanted to find the last prime factors of the integer $2^{953} - 1$
(see ``Wanted List'' of the Cunningham Project\index{Cunningham project}
\href{http://www.cerias.purdue.edu/homes/ssw/cun/}{\texttt{http://www.cerias.purdue.edu/homes/ssw/cun/}}).

The 6 smaller prime factors, already found before have been:
$$
\begin{array}{c}
3, 1907, 425796183929, \\
1624700279478894385598779655842584377, \\
3802306738549441324432139091271828121 \quad {\rm and} \\
128064886830166671444802576129115872060027.
\end{array}
$$
The first 3 factors can be easily computed\footnote{%
e.g. using CrypTool\index{CrypTool} via menu 
{\bf Indiv. Procedures \textbackslash{} RSA Demonstration \textbackslash{} 
Factorisation of a Number}.
}.
The next three prime factors were found by P.~Zimmerman
(\href{http://www.loria.fr/~zimmerma/ecmnet}{\tt http://www.loria.fr/\~{}zimmerma/ecmnet}), 
T.~Grandlund (\href{http://www.swox.se/gmp/}{\tt http://www.swox.se/gmp/})
and R. Harley during the years 1999 and 2000 using the elliptic curve factorisation method.

@@

As you notice the factorized compound numbers built of 2 prime factors are
much smaller than the especially structured numbers, for which primality 
tests\index{Primality testing}} are able to decide whether they are prime 
or not (see chapter 2.4 and 2.5).

>>>>

% --------------------------------------------------------------------------
\subsubsection{Patterns and structures}
\hypertarget{kap_3_5_2}
In general mathematicians investigate ``Structures''\index{Structure}. They ask e.g. at $ a * x \equiv b $ mod $m$,
which values $x$ can take for given values of $a, ~b, ~m.$

@@

% --------------------------------------------------------------------------
\subsubsection{Patterns and structures}
\hypertarget{Kap_3_8}{}

@@

This record, established on January 18, 2002 by researchers of the German
university of Bonn\footnote{%
\href{http://www.ercim.org/publication/Ercim\_News/enw49/franke.html, 2002-01}
{\texttt{http://www.ercim.org/publication/Ercim\_News/enw49/franke.html, 2002-01}}}, got much less attention within the press than the
solution of the RSA challenge, established on August 22, 1999, when
researchers from the Netherlands factorized the 155-digit number into its
both 78-digit primes (see Chapter \ref{chptSecurityParam}). 
One reason is that 512 bit RSA-155 was a kind of {\em magic} border.

The task of the researchers from Bonn was not initiated by a challenge, but
they wanted to find the last prime factors of the integer $2^{953} - 1$
(see ``Wanted List'' of the Cunningham Project\index{Cunningham project}%
\footnote{\href{http://www.cerias.purdue.edu/homes/ssw/cun/}{\texttt{http://www.cerias.purdue.edu/homes/ssw/cun/}}}).

The 6 smaller prime factors, already found before have been:
$$
\begin{array}{c}
3, 1907, 425796183929, \\
1624700279478894385598779655842584377, \\
3802306738549441324432139091271828121 \quad {\rm and} \\
128064886830166671444802576129115872060027.
\end{array}
$$
The first 3 factors can be easily computed\footnote{%
e.g. using CrypTool\index{CrypTool} via menu 
{\bf Indiv. Procedures \textbackslash{} RSA Demonstration \textbackslash{} 
Factorisation of a Number}.
}.
The next three prime factors were found by P.~Zimmerman
\footnote{\href{http://www.loria.fr/~zimmerma/ecmnet}{\tt http://www.loria.fr/\~{}zimmerma/ecmnet}}, 
T.~Grandlund\footnote{\href{http://www.swox.se/gmp/}{\tt http://www.swox.se/gmp/}}
and R. Harley during the years 1999 and 2000 using the elliptic curve factorisation method.

@@

As you notice the factorized compound numbers built of 2 prime factors are
much smaller than the especially structured numbers, for which primality 
tests\index{Primality testing} are able to decide whether they are prime 
or not (see chapter 2.4 and 2.5).

### 2003-03-31 script/en/chapters/primes.tex
be update fixes

\\
Despite Fermat was wrong with this supposition, he is the originator of
an important theorem in this area: Many rapid prime number tests are 
based on the (little) Fermat theorem put forward by Fermat in 1640
(\hyperlink{KleinerSatzFermat-chap3}{see chapter 3.8.3}).

@@

The following enumeration contains the most common ideas for 
``prime number formulae'':
\begin{enumerate}

\item {\em Mersenne numbers}  $f(n) = 2^n - 1$  \quad for $ n $ prime: \\
    As shown \hyperlink{MersenneNumbers01}{above}, this formula seems
    to deliver relatively large prime numbers but - as for $n=11$ 
    [$f(n)=2,047$] - it is repeatedly the case that the result even with
    prime exponents is \hyperlink{Mer-nums-not-always-prim}{not} prime.\\
    Today, all the Mersenne primes having less than around 2,000,000 digits are 
    known (M-38\index{M-38}) \\
    ({\href{http://perso.wanadoo.fr/yves.gallot/primes/index.html}
           {\tt http://perso.wanadoo.fr/yves.gallot/primes/index.html}}).

\item $F(k,n) = k \cdot 2^n \pm 1$  for $ n $ prime and $ k $ small primes:\\
   For this generalisation of the Mersenne 
   numbers\index{Mersenne!number!generalized}
   there are (for small $k$)
   also extremely quick prime number tests (see \cite{Knuth1981}). This can
   be performed in practice using software such as the Proths software from
   Yves Gallot\index{Gallot Yves} \\
   ({\href{http://www.prothsearch.net/index.html}
    {\tt http://www.prothsearch.net/index.html}}).

@@

    Within the project ``Distributed Search for Fermat Number Dividers''
    offered by Leonid Durman there is also progress in finding new monster
    primes
    ({\href{http://www.fermatsearch.org/}
       {\tt http://www.fermatsearch.org/}}  --  This website has conjectured
    webpages in Russian, Italian and German).

    The discovered factors can be compound integers or primes.

>>>>

\vspace{12pt}
Despite Fermat was wrong with this supposition, he is the originator of
an important theorem in this area: Many rapid prime number tests are 
based on the (little) Fermat theorem put forward by Fermat in 1640
(\hyperlink{KleinerSatzFermat-chap3}{see chapter 3.8.3}).

@@

The following enumeration contains the most common ideas for 
``prime number formulae'':
\begin{enumerate}
\item {\em Mersenne numbers}  $f(n) = 2^n - 1$  \quad for $ n $ prime: \\
    As shown \hyperlink{MersenneNumbers01}{above}, this formula seems
    to deliver relatively large prime numbers but - as for $n=11$ 
    [$f(n)=2,047$] - it is repeatedly the case that the result even with
    prime exponents is \hyperlink{Mer-nums-not-always-prim}{not} prime.\\
    Today, all the Mersenne primes having less than around 2,000,000 digits are 
    known (M-38\index{M-38}):
\vspace{-10pt}
\begin{itemize}
  \item[] {\href{http://perso.wanadoo.fr/yves.gallot/primes/index.html}
           {\tt http://perso.wanadoo.fr/yves.gallot/primes/index.html}}
\end{itemize}
\item $F(k,n) = k \cdot 2^n \pm 1$  for $ n $ prime and $ k $ small primes:\\
   For this generalisation of the Mersenne 
   numbers\index{Mersenne!number!generalized}
   there are (for small $k$)
   also extremely quick prime number tests (see \cite{Knuth1981}). This can
   be performed in practice using software such as the Proths software from
   Yves Gallot\index{Gallot Yves}
\vspace{-10pt}
\begin{itemize}
  \item[] {\href{http://www.prothsearch.net/index.html}
    {\tt http://www.prothsearch.net/index.html}}.
\end{itemize}

@@

    Within the project ``Distributed Search for Fermat Number Dividers''
    offered by Leonid Durman there is also progress in finding new monster
    primes: 
\vspace{-10pt}
\begin{itemize}
  \item[] {\href{http://www.fermatsearch.org/}
       {\tt http://www.fermatsearch.org/}}\\
       This website links to other webpages in Russian, Italian and German.
\end{itemize}
    The discovered factors can be compound integers or primes.
